### 2015-09-24 p1/report/latex/draft.tex
Commit

\documentclass[11pt,english]{article}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{morefloats}
\usepackage[toc,page]{appendix}
\usepackage{titlesec}
\graphicspath{ {../} }

\newcommand{\sectionbreak}{\clearpage}

\author{Sean Curran \\
\and Stephen Durofchalk \\
\and Zachary Feldman \\
\and Ryan Wails}

\title{EE454 Project 1 Report}
\lstset{language=Matlab}
\renewcommand{\thesection}{\alph{section}}

\begin{document}
\maketitle

\newpage
\section{Project Summary}

\section{Procedural Approach}

\section{Experimental Observations}

\subsection{Intermediate Output}

Intermediate output generated by the CNN is included in the appendix attached to this report.  Each figure shows the concatenated output of each layer as a gray scale image.

\section{Performance Evaluation}
\subsection{Results}
	The following results were the outcome of classifying images as the class associated with the highest probability in the probability vector.
\begin{center}
	\includegraphics[scale=0.7]{confusionmatrix}
	~\\~\\

		Accuracy = 43.7\%
\end{center}

\begin{center}
	\includegraphics[scale=0.6]{confsurf}
	~\\~\\
\end{center}
\textbf{Discussion:}

   The surface plot above demonstrates how well the CNN classifies images. There is a very strong response on the diagonal which represents the correct classifications. What is interesting is that this matrix is somewhat diagonal. This demonstrates that if images in a class of images $C_1$ are being confused for images in a class of images $C_2$, then images in $C_2$ are also being confused for images in $C_1$. A lot of this confusion is entirely logical; for example, trucks and automobiles are frequently confused for each other. Another great example of confusion are cats and dogs are frequently confused for one another. We are again very impressed with how well the CNN because it not only detects similarities between images within a class, but it detects similarities amongst images spanning multiple classes.

    ~\\
    \noindent
    Note: The above results were produced using NeuralNet function located in 'NeuralNet.m'.

\subsection{Looking at the Top-k Highest Probabilities}
	The following subsection looks at how frequently an image's true class appears in one of the top-k ranked classes, as determined by probability scores sorted from highest to lowest.
\begin{center}
	\includegraphics[scale=0.8]{topktable}

\end{center}
	Note: Accuracy is the percentage of times that the correct object class appeared in the top-k ranked classes, as determined by probability scores sorted form highest to lowest. 
	~\\
\begin{center}
	\includegraphics[scale=0.8]{topkplot}
	~\\~\\
\end{center}
\textbf{Discussion:}
      We were very surprised to see the top-k classification plot follow a logarithmic distribution. This demonstrates that the probability estimations provided by the CNN are accurate. We expected the probability estimations to be an arbitrary metric, used to find reasonable classification for an image, but looking at the logarithmic nature of the plot it is clear that the probabilities are very accurate. In fact, when looking at the probabilities for the 10,000 images you will find that on average an images top-k probabilities sum up to the y value shown above. For example, the average max probabilities for the 10,00 images is .4425 and 43.71\% of the time the max probability is correct. Even more interesting, the average of the top 2 probabilities added together for each image is .6646 and the chance of the true class being associated with one of the top two highest probabilities in its probability vector .6591, as shown above.

	~\\
	\noindent
	Note: The above results were produced using the topk function located in 'topk.m' and NeuralNet function located in 'NeuralNet.m'.

\section{Further Exploration}
\subsection{Test Cases}
To test the robustness/accuracy of the training outside of the given data set, additional test images were gathered and fed into the CNN.  Images were rescaled to thumbnail size (32x32x3) with the Matlab command
\begin{lstlisting}
thumbnail = imresize(img, [32 32]);
\end{lstlisting}

\noindent
110 images total were collected; 100 fell into the preexisting image classes, 10 contained scenes not falling into any preexisting class.  Running these new 10 images through the CNN yielded the following results:\\

\begin{tabular}{ | c | c | c |}
  \hline
  Image \# & Image Contents & CNN Classification \\
  \hline		
  Image 1 & Tree in Field & Horse (8) \\
  Image 2 & Winter Mountain Scene & Ship (9) \\
  Image 3 & Ryan and Zach & Bird (3) \\
  Image 4 & Steve & Frog (7) \\
  Image 5 & Sean & Cat (4) \\
  Image 6 & Desktop Computer & Automobile (2) \\
  Image 7 & Football on Field & Deer(5) \\
  Image 8 & Robert Collins & Truck (10) \\
  Image 9 & Old Main & Bird(3) \\
  Image 10 & Penn State Logo & Truck (10)\\
  \hline  
\end{tabular}

~\\~\\
\noindent
All of the exploratory images we used were either taken by us or were taken from Google Images. The code we used to read in the exploratory images can be found in 'getAllFiles.m' and 'getTrueClass.m'. We produced the results above by running the 10 images through the function found in 'NeuralNet.m' and classifying the images by assigning the image the class associated with the highest probability. The classification process was done manually.

\subsection{Reclassification}
An attempt was made to distinguish these 10 images from the rest of the data set (i.e. reclassify these 10 images as unknown).  Let $V$ be the vector of output probabilities from the CNN for each image.  Let $C$ be the classification of the image producing probability vector $V$.  Originally, we have\\~\\
\begin{math}
p_i \in V \\
p_{max} = \max(p_1, p_2,...,p_{10}) \\
C = i $ given $ p_i = p_{max}
\end{math}

~\newline\noindent
To reclassify the images we use\\
\noindent
\begin{math}
p_{max} = \max(p_1,p_2,...,p_{10}) \\
P = \{p_1,p_2,...,p_{10}\} \setminus \{p_{max}\} \\
p_{2max} = \max(P) \\~\\
\begin{cases}
C = i $ given $ p_i = p_{max} & $for	$ p_{max} - p_{2max} >= 0.25 \\
C = 11 & $otherwise$
\end{cases}
\end{math}
~\\~\\
Less precisely, the difference between the two peak responses was thresholded by .25; so, if there were two strong responses, the image is reclassified as unknown.
~\\~\\
The code used to reclassify images can be found in 'identify\_exploratory.m'.
\subsection{Results After Reclassification}
In the confusion matrix below, classes 1 through 10 are the same classifications.  Class 11 identifies an unknown image class. \\\\
\begin{tabular}{ | c | c | c | c | c | c | c | c | c | c | c | c | c | }
\hline
	 & CNN
Class & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 \\ \hline
	True 
Class &  &  &  &  &  &  &  &  &  &  &  &  \\ \hline
	1 &  & 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 6 \\ \hline
	2 &  & 1 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 0 & 5 \\ \hline
	3 &  & 1 & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 7 \\ \hline
	4 &  & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 9 \\ \hline
	5 &  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 9 \\ \hline
	6 &  & 0 & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 8 \\ \hline
	7 &  & 0 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 0 & 8 \\ \hline
	8 &  & 0 & 0 & 1 & 1 & 0 & 0 & 1 & 4 & 0 & 0 & 3 \\ \hline
	9 &  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 5 & 0 & 5 \\ \hline
	10 &  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 5 & 5 \\ \hline
	11 &  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 9 \\ \hline
\end{tabular}

\includegraphics[scale=0.6]{confexp}
~\\~\\
\textbf{Accuracy = 29.09\%}
\\
The reclassification metric does well in replacing the 10 new images in the unknown category, but also places images that were previously correctly classified in the unknown category.  Other metrics tested were
\begin{enumerate}
\item Thresholding on the number of classes that had a response above 0.1 (the mean value for random classification)
\item Taking the spatial derivative of the probability vector; this metric actually produces zero-mean Gaussian noise with a very small standard deviation.
\end{enumerate}

~\\~\\
The code used to produce the results found in this section can be found in 'exploratory\_main.m'.

\section{Member Contributions}
	Overall the contributions made form each team member were pretty even. Whenever the team met to work on the project most of the team members were in attendance. Everyone who was in attendance at each team meeting attempted to work on a different piece of the project to ensure that we were making steady progress. The coding was not split up evenly as one team member did not attend the initial meeting and as a result he missed most of the actual coding and basically all of the design decisions. We attempted to split the report up evenly, but as the deadline approached scheduling meetings became difficult and some individuals were unavailable to work on the report. In conclusion, the work was not split up evenly among our four group members, but everyone attempted to contribute.

\newpage
\begin{appendices}
\section{Intermediate Output}

\begin{figure}[h!]
  \caption{Original Image}
  \centering
    \includegraphics[width=0.75\textwidth]{layer/0}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 1 Output}
  \centering
    \includegraphics[width=0.75\textwidth]{layer/1}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 2 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/2}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 3 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/3}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 4 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/4}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 5 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/5}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 6 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/6}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 7 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/7}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 8 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/8}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 9 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/9}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 10 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/10}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 11 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/11}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 12 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/12}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 13 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/13}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 14 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/14}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 15 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/15}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 16 Output}
  \centering
    \includegraphics[width=\textwidth]{layer/16}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 17 Output}
  \centering
    \includegraphics[width=0.75\textwidth]{layer/17}
\end{figure}

\begin{figure}[h!]
  \caption{Layer 18 Output}
  \centering
    \includegraphics[width=0.75\textwidth]{layer/18}
\end{figure}

\end{appendices}

\end{document}