### 2015-12-11 p2/report/project_report.tex
finished harris code

\documentclass{article}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{draftwatermark}
\usepackage {sectsty}
\usepackage{hyperref}
\usepackage [dvipsnames]{xcolor}

\lstset{ %
	language=C,
	basicstyle=\ttfamily\footnotesize,
	keywordstyle=\ttfamily\color{blue}\footnotesize,
	commentstyle=\color{OliveGreen}\ttfamily\footnotesize,
	texcl,
	escapechar =`
	escapebegin=\lst@commentstyle,
	showspaces=false,
	showtabs=false,
	numbers=left,
	showstringspaces=false,
	frame=single
}
\makeatother

\pagestyle{fancy}
\renewcommand{\rightmark}{}

\SetWatermarkText{DRAFT}
\SetWatermarkScale{5}
\SetWatermarkLightness{.98}

\sectionfont{\color{cyan}}
\author{Sean Curran \\
\and Stephen Durofchalk \\
\and Zachary Feldman \\
\and Ryan Wails}
\title{CMPEN/EE 454 Project 2 Report}

\begin{document}

\maketitle
\thispagestyle{empty}

\newpage
\pagenumbering{roman}
\setcounter{page}{1}
\setcounter{tocdepth}{2}
\tableofcontents
\thispagestyle{plain}

\newpage
\pagenumbering{arabic}
\setcounter{page}{1}
\section{Integral Algorithms}
\subsection{Harris Corner Detection}
\subsubsection{Synopsis}
Harris corner detection is an algorithm which detect corners in images that can be used to create point correspondences between time series images.  A ``corner'' can loosely be defined as a feature in the image with relatively large $\frac{\partial f}{\partial x}$ and $\frac{\partial f}{\partial y}$ values.
\subsubsection{Implementation}
Below the implementation of the Harris corner detection algorithm in C-like pseudocode.
\begin{lstlisting}

intensity_image harris(original_image, patch_indices) {

	// Compute the image patch.
	intensity_img = convert_to_intensity_image(original_image);
	img_patch = compute_patch(intensity_img, patch_indices);

	// Compute partial x and y, mixed derivatives.
	ix = convolve_sobel_x(img_patch);
	iy = convolve_sobel_y(img_patch);

	ix2 = ix * ix;
	iy2 = iy * iy;
	ixy = ix * iy;

	// Compute sums using a Gaussian windowing function
	sx2 = convolve_gaussian(ix2);
	sy2 = convolve_gaussian(iy2);
	sxy = convolve_gaussian(ixy);

	// For each pixel in the image, compute H matrix,
	// compute product of eigenvalues scaled by k = 0.07,
	// and store the response value as r.
	for (px : img_patch) {
		H = {{sx2(px), sxy(px)}, {sxy(px) sy2(px)}};
		r(px) = det(H) - 0.07 * trace(H) * trace(H);
	}

	return r;
}

\end{lstlisting}
\subsection{Patch Matching}
\subsection{RANSAC}

\end{document}